Transient Fault Tolerant Asynchronous Circuits

### 2019-12-06 draft_final.tex
before final ch 2 changes

Asynchronous circuits differ from synchronous circuits in that the gates have arbitrary delays so asynchronous designs must be more meticulous to account for these delays.  The arbitrary delays also make fault-tolerance a harder problem.  For example, the classic solution of Triple Modular Redundancy (TMR) does not work for asynchronous circuits.  The main contribution of this work is proposing a new fault-tolerant design for asynchronous circuits.  The new design expands fault-tolerance to semi-modular speed independent circuits composed of two-input gates and C-elements.  This work also proves that the proposed fault-tolerant design is tolerant to transient faults.  With this new design, circuits can be more robust to different types of errors, allowing the circuits to be used in a greater variety of environments.

@@

\section{Synchronous vs Asynchronous circuits}
Digital circuits can be divided into two types, synchronous and asynchronous.  In digital circuits, the time it takes for gates to process input signals into output signals and for these signals to reach the next gates vary widely.  If the timing is somehow off, it is possible for the outputs to be different from the desired value resulting in incorrect computation. In synchronous systems this timing variation is dealt with by using a global clock.  The clock is then fed into each flip flop so that the flip flop outputs will lock and update the signals synchronously. The clock cycle frequency is chosen so that all of the gates can finish processing and the correct values are locked for the next cycle. That is, the clock cycle is usually designed to accommodate the slowest logic element. The design of synchronous systems can be complex and is not necessary to understand in detail for this work. \\
\begin{figure}
\includegraphics[width=\textwidth]{syncex}
\captionof{figure}[Example of a synchronous digital circuit]{Example of a synchronous digital circuit.  Each output is gated by a flip flop that is controlled by the clock}
\end{figure}
%possible picture of clock and flip flops?.... redraw as my own

@@

The asynchronous delay model that we consider is the unbounded gate delay model.  Circuits that can function correctly under arbitrary gate delays are called Speed Independent circuits. The unbounded gate delay model assumes that delays only appear on the gates and the delays can be arbitrarily long in the interval $(0,\infty)$.  The wire delays are negligible.  In addition we assume all forks, which is splitting a signal into multiple branches, do not have any time delays between each of its branches.  These assumptions are reasonable because the delays on the wires only change the operation of a circuit if the wire has forks.  (Otherwise you can lump the wire delay with the previous gate)  When a wire has forks, it is usually true in implementation that the delay between different branches of the fork is less than the delay for gates so that they arrive at the next gates at the same time from the perspective of the relevant gates.  This is also termed isochronic forks.  We focus on the unbounded gate delay model because the delay being unbounded is a pretty harsh requirement and would allow more circuits to work in a real life setting. \\ %how if we assume both gates and wires have delay cannot create many SI circuits, and also Muller is one of the broader classes that we can use

D. E. Muller proposed essentially this class of speed independent circuits, which are sometimes called Muller circuits \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synthesis methods begin with some higher-level specification such as a signal transition graph or Petri net and translate into a state graph.  A state graph describes all the possible sequences of changes in a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allow the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed independent asynchronous circuit.  The construction we describe begins with a semi-modular circuit, possibly resulting from the previously described process, and transforms it into a circuit that can tolerate any single transient fault.

@@

\section{Organization of this work}
%why are async circuits important, timing delay etc
Because these various faults can interfere with the function of the circuit and even result in harmful operations (catastrophic failures), there has been a lot of work on making digital circuits fault-tolerant. Most of the fault-tolerant solutions have been investigated for synchronous digital circuits.  Because the timing difference between different signals arriving do not have to be considered for synchronous circuits, they can use error correcting codes and redundancy techniques such as triple modular redundancy (TMR). Unfortunately most of these techniques are not applicable to asynchronous circuits.\\

@@

\chapter{Related Works}
\section{Fault Model}
A fault is defined as the physical difference between the good or correct system, and the current system.  In this work, the system is the asynchronous circuit.  An error, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function, may be caused by a fault.  Hidden faults, where the fault does not cause an error, are also possible \cite{jha_gupta_2003}. \\

Faults can be split into permanent and non-permanent faults. Permanent faults are faults that affect the functional behavior of a system or circuit permanently.  Non-permanent faults are faults that affect the system's functional behavior for a finite, but unknown period of time.  Non-permanent faults can occur at random moments and are categorized into two types:  transient faults and intermittent faults \cite{jha_gupta_2003}.\\ %copied

@@

\subsection{Speed Independent Circuits}
Circuits are composed of logic elements and wires. The unbounded gate delay model assumes an arbitrary delay on the logic elements and no delay on wires.  This delay model is the basis for Muller circuits \cite{Muller_59}.  Circuits that operate correctly regardless of the delay are called Speed Independent Circuits. The designer specifies the correct operation of the circuit.  For example, given certain input values, the circuit will return specified output values. 
There are many ways to represent specifications such as with Petri nets and state graphs, and there are established methods to build asynchronous circuits that satisfy the specifications \cite{myers_book_2004}. We won't go into these representations and methods in detail, other than the aspects that are relevant to this work. We work with digital circuits, so the wires take on Boolean values of either `1' or `0'.\\

@@

We consider semi-modular circuits because they are a subset of speed independent circuits and is a large enough class to include some interesting and useable circuits. A simple explanation of semi-modularity is that once a wire is excited, it stays excited until it transitions to the excited value. Formally this can be defined as follows: %proof that semi-modular circuits are also speed independent
\begin{definition}An asynchronous circuit is {\em semi-modular} if for every pair of states $s$ and $s'$ such that $T(s,s')$ holds, and let $v$ and $w$ be wires such that $s(w) \neq s'(w)$ and $w\neq v$, then if $v$ is excited in state $s$, $v$ is also excited in state $s'$ \end{definition}

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph has the following properties, it is possible to synthesize a hazard free and speed independent asynchronous circuit following the state graph. Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire. We next look at how such a gate logic can be made.\\ 
\end{comment}

@@

The standard C-implementation is a gate-level implementation.  Each wire from the specification is implemented as a C-element with non-looping cascades of gates as the C-element's two inputs.  That is, each of the two inputs to the C-element is itself an output of a cascade of gates.  The C-element operates like a typical C-element where if the two inputs are `1', then the output is `1'; and when the two inputs are `0', the output is `0'; and when the two inputs have different values, the output value is held.  Additionally, Myers shows a method where the gate logic for the inputs to the C-element can always be re-implemented with two-input gates \cite{myers_book_2004}.  Since the C-element itself is also a two-input circuit element, then the entire circuit is composed of two-input circuit elements.  In such an implementation, each gate or C-element is assumed to have its own delay that is arbitrary in length.  Thus, this gate-level implementation allows for more generic design.
%explain sop?
\begin{comment}
\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

Another important property of QDI circuits is \textit{stability}. A guard is stable if it evaluates to False after the transition of the output variable is complete. This eliminates any hazards that may arise when a guard is True and before the output has fully transitioned, the guard then evaluates to False and the transition is terminated. This may cause a blip in the voltage, an incomplete transition, where the circuit may interpret it as either no transition or two consecutive transitions. One additional assumption is that the PRs cannot be \textit{self-invalidating}. That is, when a transition on an output occurs, it cannot immediately falsify the guard. With this assumption, it can be shown that the output of a gate cannot also be its own input variable. There are no self-loops. 
\end{comment}
\section{Fault Tolerant solutions}

@@

\chapter{Asynchronous Circuit Concepts and Definitions}
\section{Assumptions:  Boolean signals}
This work assumes that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the gates and C-elements are excited and their outputs transition, we assume that the transition is instantaneous and that, after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

@@

\section{Circuit Elements}
Gates and elements (C-elements) are components that we allow in our circuit.  In general, we call them {\em elements}.  Each element has two input wires and one output wire.  The elements can be in two states, excited or stable.  Examples of gates are AND, OR, NOT etc.

For each element, there exists a function that maps the inputs into a binary value.  This function is represented by $f:[I \to \{0,1\}] \to \{0,1\}$ where $I$ is the set of input wires of the element, a subset of $W$.  If we assume a two input element, we name the input wires $w_1$ and $w_2$.  The function models the designed operation of the element such that if the element had no delay, $f(w_1,w_2)$ is the output value of the element.  With asynchronous delay, if the inputs remain constant, then the output will eventually take the value of $f(w_1,w_2)$.  For gates, such a function is easily defined.  For example, an AND gate such as in Figure \ref{fig:gates0} with two inputs $w_1$ and $w_2$ is modeled by the function 
\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{gates0}
\captionof{figure}{AND gate}
\label{fig:gates0}
\end{figure}
\[
f_{AND}(w_1,w_2,v)= w_1\wedge w_2
\]

For C-elements, these are memory holding devices so the function also requires the previous output.  For example, the C-element in Figure \ref{fig:element0} with two input wires $w_1$ and $w_2$ is modeled by the function\\
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{element0}
\captionof{figure}{C-element}
\label{fig:element0}
\end{figure}
\[
f_{C-E}(w_1,w_2, v) =\left\{
  \begin{array}{@{}ll@{}}
    0, & \text{if}\ w_1=0 \wedge w_2=0\\
    1, & \text{if}\ w_1=1 \wedge w_2=1\\
    v, & \text{otherwise}

  \end{array}\right. 
\]
When $f_{C-E}(w_1,w_2, v) =v$, we say the C-element is in the ``hold state''.  
Each element is in one of two states, excited or stable.  
\begin{definition}
An element with inputs $w_1$, $w_2$, output $v$ and function $f$ is {\em excited} if $v\neq f(w_1,w_2, v)$.  
\end{definition}

@@

When the element is excited, there are two possibilities for the element to become stable again.  Either the input changes such that the element is stable, or the output changes.  Again, not all input changes will cause the element to become stable.  In fact, semi-modular circuits utilizes input changes that allow the element to stay excited.  If the output changes in an excited state, the element becomes stable and we call this action a {\em firing} of the element.

\section{Circuits}
A circuit is a network of elements.  These elements are connected when they share a wire.  For an element, each input is connected to either a primary input or the output of a different element.  Other than elements, a circuit also consists of primary inputs and primary outputs.  Primary inputs are all inputs that are not the output of an element in the circuit.  Primary outputs are outputs of an element in the circuit that connects to another device or element outside of the circuit.  Primary inputs and outputs can be thought of as the interface with outside of the circuit.  They can be the outputs and inputs of a different circuit, or connected to physical mechanisms such as a user pushing a button.   An example of a circuit with primary inputs and outputs is:  .

\section{States}
Let $W$ be the set denoting the union of all wires of elements and the primary inputs of a circuit, then $S$ is the set of functions $[W\to \{0,1\}]$ which maps each element of $W$ to a boolean value.  An element of $S$ is called a {\em state}.

Assigning values to all wires of a circuit names a state.  An example of a state for a circuit containing 3 wires is `101'.  From the set definition, states are unique and no two states share the same values.

\section{Behavior of a circuit}    
Circuits respond to input changes and produces outputs that accomplish some specified task.  Thus the function of a circuit depends on its dynamic behavior and how its state changes over time.  Under the state representation of a circuit, if the value of a wire changes, the state also changes.  If the wire changes from a previous state $s$ to a new state $s'$, then there is a transition relation between the two states.  Defined more formally with $s(w)$ denoting the value of wire $w$ in state $s$:  
\begin{definition}
For all $s$ and $s'$ in S, the \emph{transition relation} $T(s,s')$ holds when there is exactly one wire $w \in W$ such that $s(w)\neq s'(w)$, and wire $w$ must be either the output of an excited element in state $s$, or $w$ is a primary input.
\end{definition}
%from state graph, T(s,s') must satisfy... but we can define ourselves which are the transitions (ie arrows in a state graph).  Going from the circuit, there is a transition if the circuit state allows for a transition, either something excited fires or the primary input changes
There may be multiple transitions from a state $s$.

We define a {\em trace} to describe a path the circuit can take.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states the trace, which is formally defined as follows:
\begin{definition}A {\em trace} $\sigma=(s_1, s_2, ...)$ of a circuit is a sequence of states of finite or infinite length.  
%Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Every pair of consecutive states in a trace are in a transition relation $T(s_i,s_{i+1})$, where state $s_i$ represents the $i^{th}$ term of the sequence.  .  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
\end{definition} 
We also use $\sigma_n$ as shorthand to denote the sequence of the first $n$ states of $\sigma$.
When there multiple transitions from a state, multiple traces exist depending on which wire transitions.  Thus we want to define a set of traces for all the traces that form.  In addition the initial state is important in naming a trace and by extension for a set of traces.  The {\em initial state} specifies the state of the wires in the circuit as well as the primary inputs.  Thus a set of traces is defined as follows:
\begin{definition}
A {\em set of traces} of a circuit with initial state $s_1$ is the set of all traces of the circuit starting in state $s_1$. 
\end{definition}

 % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
 The circuit may be waiting for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
%\begin{definition}
%A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
%\end{definition}
%n other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces can describe a circuit completely, in addition we may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state $s$ is {\em reachable} from $s_1$ if $s$ appears in some trace that begins with $s_1$.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces. 
%still kind of weird oh well 
\end{definition}
Thus, unreachable states are states that are not in any of the traces with initial state $s_1$.  Since all functionality of a circuit can be described by a set of traces starting in a specified state, then the circuit will never enter those unreachable states.

>>>>

Asynchronous circuits differ from synchronous circuits in that the gates have arbitrary delays so asynchronous designs must be more meticulous to account for these delays.  The arbitrary delays also make fault-tolerance a harder problem.  For example, the classic solution of Triple Modular Redundancy (TMR) does not work for asynchronous circuits.  The main contribution of this work is proposing a new fault-tolerant design for asynchronous circuits.  The new design expands fault-tolerance to semi-modular speed-independent circuits composed of two-input gates and C-elements.  This work also proves that the proposed fault-tolerant design is tolerant to transient faults.  With this new design, circuits can be more robust to different types of errors, allowing the circuits to be used in a greater variety of environments.

@@

\section{Synchronous vs Asynchronous Circuits}
Digital circuits can be divided into two types: synchronous and asynchronous.  In digital circuits, the time it takes for gates to process input signals into output signals and for these signals to reach the next gates vary widely.  If the timing is somehow off, it is possible for the outputs to be different from the desired value resulting in incorrect computation. In synchronous systems this timing variation is dealt with by using a global clock.  The clock is then fed into each flip flop so that the flip flop outputs will lock and update the signals synchronously. The clock cycle frequency is chosen so that all of the gates can finish processing and the correct values are locked for the next cycle. That is, the clock cycle is usually designed to accommodate the slowest logic element. The design of synchronous systems can be complex and is not necessary to understand in detail for this work. \\
\begin{figure}
\includegraphics[width=\textwidth]{syncex}
\captionof{figure}[Example of a synchronous digital circuit]{Example of a synchronous digital circuit.  Each output is gated by a flip flop that is controlled by the clock}
\end{figure}
%possible picture of clock and flip flops?.... redraw as my own

@@

The asynchronous delay model that we consider is the unbounded gate delay model.  Circuits that can function correctly under arbitrary gate delays are called speed-independent circuits. The unbounded gate delay model assumes that delays only appear on the gates and the delays can be arbitrarily long in the interval $(0,\infty)$.  The wire delays are negligible.  In addition we assume all forks, which is splitting a signal into multiple branches, do not have any time delays between each of its branches.  These assumptions are reasonable because the delays on the wires only change the operation of a circuit if the wire has forks.  (Otherwise you can lump the wire delay with the previous gate)  When a wire has forks, it is usually true in implementation that the delay between different branches of the fork is less than the delay for gates so that they arrive at the next gates at the same time from the perspective of the relevant gates.  This is also termed isochronic forks.  We focus on the unbounded gate delay model because the delay being unbounded is a pretty harsh requirement and would allow more circuits to work in a real life setting. \\ %how if we assume both gates and wires have delay cannot create many SI circuits, and also Muller is one of the broader classes that we can use

D. E. Muller proposed essentially this class of speed-independent circuits, which are sometimes called Muller circuits \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synthesis methods begin with some higher-level specification such as a signal transition graph or Petri net and translate into a state graph.  A state graph describes all the possible sequences of changes in a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allow the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed-independent asynchronous circuit.  The construction we describe begins with a semi-modular circuit, possibly resulting from the previously described process, and transforms it into a circuit that can tolerate any single transient fault.

@@

\section{Organization of This Work}
%why are async circuits important, timing delay etc
Because these various faults can interfere with the function of the circuit and even result in harmful operations (catastrophic failures), there has been a lot of work on making digital circuits fault-tolerant. Most of the fault-tolerant solutions have been investigated for synchronous digital circuits.  Because the timing difference between different signals arriving do not have to be considered for synchronous circuits, they can use error correcting codes and redundancy techniques such as triple modular redundancy (TMR). Unfortunately most of these techniques are not applicable to asynchronous circuits.\\

@@

\chapter{Related Work}
\section{Fault Model}
A fault is defined as the physical difference between the good or correct system, and the current system.  In this work, the system is the asynchronous circuit.  An error, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function, may be caused by a fault.  Hidden faults, where the fault does not cause an error, are also possible \cite{jha_gupta_2003}. \\

Faults can be classified as permanent and non-permanent faults. Permanent faults are faults that affect the functional behavior of a system or circuit permanently.  Non-permanent faults are faults that affect the system's functional behavior for a finite, but unknown period of time.  Non-permanent faults can occur at random moments and are categorized into two types:  transient faults and intermittent faults \cite{jha_gupta_2003}.\\ %copied

@@

\subsection{Speed-independent Circuits}
Circuits are composed of logic elements and wires. The unbounded gate delay model assumes an arbitrary delay on the logic elements and no delay on wires.  This delay model is the basis for Muller circuits \cite{Muller_59}.  Circuits that operate correctly regardless of the delay are called speed-independent circuits. The designer specifies the correct operation of the circuit.  For example, given certain input values, the circuit will return specified output values. 
There are many ways to represent specifications such as with Petri nets and state graphs, and there are established methods to build asynchronous circuits that satisfy the specifications \cite{myers_book_2004}. We won't go into these representations and methods in detail, other than the aspects that are relevant to this work. We work with digital circuits, so the wires take on Boolean values of either `1' or `0'.\\

@@

We consider semi-modular circuits because they are a subset of speed-independent circuits and is a large enough class to include some interesting and useable circuits. A simple explanation of semi-modularity is that once a wire is excited, it stays excited until it transitions to the excited value. Formally this can be defined as follows: %proof that semi-modular circuits are also speed independent
\begin{definition}An asynchronous circuit is {\em semi-modular} if for every pair of states $s$ and $s'$ such that $T(s,s')$ holds, and let $v$ and $w$ be wires such that $s(w) \neq s'(w)$ and $w\neq v$, then if $v$ is excited in state $s$, $v$ is also excited in state $s'$ \end{definition}

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph has the following properties, it is possible to synthesize a hazard free and speed-independent asynchronous circuit following the state graph. Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire. We next look at how such a gate logic can be made.\\ 
\end{comment}

@@

The standard C-implementation is a gate-level implementation.  Each wire from the specification is implemented as a C-element with non-looping cascades of gates as the C-element's two inputs.  That is, each of the two inputs to the C-element is itself an output of a cascade of gates.  The C-element operates like a typical C-element where if the two inputs are `1', then the output is `1'; and when the two inputs are `0', the output is `0'; and when the two inputs have different values, the output value is held.  Additionally, Myers shows a method where the gate logic for the inputs to the C-element can always be re-implemented with two-input gates \cite{myers_book_2004}.  Since the C-element itself is also a two-input circuit element, then the entire circuit is composed of two-input circuit elements.  In such an implementation, each gate or C-element is assumed to have its own delay that is arbitrary in length.  Thus, this gate-level implementation allows for more generic design.
%explain sop?
\begin{comment}
\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed-independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

Another important property of QDI circuits is \textit{stability}. A guard is stable if it evaluates to False after the transition of the output variable is complete. This eliminates any hazards that may arise when a guard is True and before the output has fully transitioned, the guard then evaluates to False and the transition is terminated. This may cause a blip in the voltage, an incomplete transition, where the circuit may interpret it as either no transition or two consecutive transitions. One additional assumption is that the PRs cannot be \textit{self-invalidating}. That is, when a transition on an output occurs, it cannot immediately falsify the guard. With this assumption, it can be shown that the output of a gate cannot also be its own input variable. There are no self-loops. 
\end{comment}
\section{Fault-tolerant Solutions}

@@

\chapter{Asynchronous Circuit Concepts and Definitions}
\section{Assumptions:  Boolean Signals}
This work assumes that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the gates and C-elements are excited and their outputs transition, we assume that the transition is instantaneous and that, after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

@@

\section{Circuit Elements}
Boolean gates and C-elements are components that are allowed in our circuit.  In general, we call them {\em elements}.  Each element has two input wires and one output wire.  The elements can be in two states, excited or stable.  Examples of gates are AND, OR, NOT etc.\\

An element is defined by a function $f$ that maps the inputs and previous output into a binary value.  We name the two input wires $w_1$ and $w_2$ and the output wire $v$.  The function models the designed operation of the element such that if the element had no delay, $f(w_1,w_2, v)$ is the output value of the element.  With asynchronous delay, if the inputs and previous output remain constant, then the output will eventually take the value of $f(w_1,w_2, v)$.  For Boolean gates, such a function is easily defined.    
For example, an AND gate such as in Figure \ref{fig:gates0} with inputs $w_1$ and $w_2$ and output $v$ is modeled by the function 
\[
f_{AND}(w_1,w_2,v)= w_1\wedge w_2
\]

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{gates0}
\captionof{figure}{AND gate}
\label{fig:gates0}
\end{figure}

For Boolean gates, $f(w_1, w_2,v)$ does not depend on $v$.  
C-elements are memory holding devices, so the function depends on the previous output $v$.  For example, the C-element in Figure \ref{fig:element0} with two input wires $w_1$ and $w_2$ is modeled by the function
\[
f_{C-E}(w_1,w_2, v) =\left\{
  \begin{array}{@{}ll@{}}
    0, & \text{if}\ w_1=0 \wedge w_2=0\\
    1, & \text{if}\ w_1=1 \wedge w_2=1\\
    v, & \text{otherwise}

  \end{array}\right. 
\]

\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{element0}
\captionof{figure}{C-element}
\label{fig:element0}
\end{figure}

When $f_{C-E}(w_1,w_2, v) =v$, we say the C-element is in the ``hold state''.  
Each element is in one of two states, excited or stable.  
\begin{definition}
An element with inputs $w_1$, $w_2$, output $v$ and function $f$ is {\em excited} if $v\neq f(w_1,w_2, v)$.  
\end{definition}

@@

When the element is excited, there are two possibilities for the element to become stable again.  Either the input changes such that the element is stable, or the output changes.  Again, not all input changes will cause the element to become stable. If the output changes in an excited state, the element becomes stable and we say that the element {\em fires}.

\section{Circuits}
A circuit is a network of elements.  These elements are connected when they share a wire.  For an element, each input is connected to either a primary input or the output of a different element.  Other than elements, a circuit also consists of primary inputs and primary outputs.  Primary inputs are inputs that are not the output of an element in the circuit.  Primary outputs are outputs of an element in the circuit that connects to another device or element outside of the circuit.  Primary inputs and outputs can be thought of as the interface with the outside of the circuit.  They can be the outputs and inputs of a different circuit, or connected to physical mechanisms such as a user pushing a button.

\section{States}
Let $W$ be the set denoting the union of all wires of elements and the primary inputs of a circuit, then $S$ is the set of functions $[W\to \{0,1\}]$ which map each element of $W$ to a boolean value.  An element of $S$ is called a {\em state}.  If $w$ is a wire and $s$ a state in $S$, then $s(w)$ is the Boolean value of the wire in that state.

\section{Behavior of a Circuit}    
Circuits respond to input changes and produce outputs that accomplish some specified task.  Thus the function of a circuit depends on its dynamic behavior and how its state changes over time.  The possible transitions between consecutive states are defined by the transition relation of a circuit.
\begin{definition}
For all $s$ and $s'$ in S, the \emph{transition relation} $T(s,s')$ holds when there is exactly one wire $w \in W$ such that $s(w)\neq s'(w)$, and wire $w$ is either the output of an excited element in state $s$, or $w$ is a primary input.
\end{definition}
%from state graph, T(s,s') must satisfy... but we can define ourselves which are the transitions (ie arrows in a state graph).  Going from the circuit, there is a transition if the circuit state allows for a transition, either something excited fires or the primary input changes
There may be multiple transitions from a state $s$.

We define a {\em trace} to describe a trajectory that the circuit can take.  Different trajectories can occur because there may be multiple alternative transitions from each state.  We call the succession of states the trace, which is formally defined as follows:
\begin{definition}A {\em trace} $\sigma=(s_1, s_2, ...)$ of a circuit is a sequence of states of finite or infinite length.  
%Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Every pair of consecutive states in a trace are in a transition relation $T(s_i,s_{i+1})$, where state $s_i$ represents the $i^{th}$ term of the sequence.  .  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
\end{definition} 
We also use $\sigma_n$ as shorthand to denote the sequence of the first $n$ states of $\sigma$.
When there are multiple transitions from a state, multiple traces exist or transitions occur.  Thus, we want to define a set of traces for all possibilities.\\

The {\em initial state} is important in defining the set of traces of a circuit.  The {\em initial state} specifies the state of the wires in the circuit as well as the primary inputs.  Thus a set of traces is defined as follows:
\begin{definition}
A {\em set of traces} of a circuit with initial state $s_1$ is the set of all traces of the circuit starting in state $s_1$. 
\end{definition}

 % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
A stopping state typically occurs when the circuit waits for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state, where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
%\begin{definition}
%A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
%\end{definition}
%n other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces describes a circuit completely. \\

We may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state $s$ is {\em reachable} from $s_1$ if $s$ appears in some trace that begins with $s_1$.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces. 
%still kind of weird oh well 
\end{definition}
Thus, unreachable states are states that do not appear in any of the traces with initial state $s_1$.  Since all functionality of a circuit can be described by a set of traces starting in a specified state, the circuit will never enter those unreachable states.

\section{Semi-modularity}
Semi-modular circuits are a subset of speed-independent circuits. 
The class of semi-modular circuits is large enough to include circuits used in real applications.  Also, they have been extensively studied, with algorithms to synthesize and analyze them.  A simple explanation of semi-modularity is that whenever an element is excited, it stays excited until it fires and its output changes.  Formally this can be defined as follows: %proof that semi-modular circuits are also speed independent
\begin{definition}
An asynchronous circuit is {\em semi-modular} if for every pair of successive states $s_{i}$ and $s_{i+1}$ in every trace, if element $j$ is excited in state $s_i$ and did not fire, then $j$ is also excited in state $s_{i+1}$. 
%An asynchronous circuit is {\em semi-modular} if for every pair of successive states in its set of traces $s_{i}$ and $s_{i+1}$, and let $v$ and $w$ be output wires such that $s_i(w) \neq s_{i+1}(w)$ (wire $w$ transitions) and $w\neq v$, then if $v$ is excited in state $s_i$, $v$ is also excited in state $s_{i+1}$ 
\end{definition}
In a circuit with primary inputs, the changes in the primary inputs are constrained not to violate semi-modularity, for example through a protocol using the primary outputs from the circuit to signal when the circuit is ready for a primary input change.
